Gut-resident microorganisms and their genes are associated with cognition and neuroanatomy in children

### 2023-10-04 manuscript/main.tex
Apply suggestions from code review
First review of PR by @kescobo

Co-authored-by: Kevin Bonham <kbonham@wellesley.edu>

\subsubsection*{Gut microbial taxa relate diffrentially to distinct cognitive subscales} %TODO: add numbers!

Is it possible to associate the predictive species with different components of cognition? 
After probing the aforedescribed importances of potential of taxa and ECs 
towards composite cognitive scores, we proceeded to verify if and how this relationship 
would hold for different subcomponents of the cognitive function assessment. 
We collected those in the form of the 
Mullen Scales of Early Learning (MSEL - \textit{see Methods}). 
Taxonomic profiles coupled with demographics exhibited predictive potential towards three probed subscales (Figure 4A), 
namely Expressive Language (EL; test-set RMSE XXX, test-set correlation YYYY), 
Gross Motor (GM; test-set RMSE XXX, test-set correlation YYYY) and 
Visual Reception (VR; test-set RMSE XXX, test-set correlation YYYY).
Between the subscales, not only the predictability (measured by the test-set figures of merit) is different,
but also the relative importances of several taxa (Figure 4B).
Although certain species such as \textit{B. longum}, \textit{F. prausznitzii} 
have similar relative importance across all subscales, others are substantially different.
Notably, \textit{Bifidobacterium pseudocatenulatum}, \textit{B. wexlerae} and \textit{E. eligens} are
substantially more important for the prediction of EL, while
\textit{R. faecis}, \textit{S. salivarius} and \textit{F. saccharivorans} are
distinctly predictive of GM, and
\textit{Clostridium innocuum}, \textit{B. vulgatus} 
stand out on models targeting VR.

@@

If there are causal effects of microbial metabolism on cognitive
function, they might be reflected in changes in neuroanatomy. We again
used RF modeling to associate gut taxonomic
profiles with individual brain regions identified in MRI scans
(children over 18 months old, N = 124).
Some brain regions were more readily
predicted by RF models trained on microbial taxa (Table 4, Table S2), in particular those that were highly correlated with age.
These included the L/R lingual gyrus (mean RF correlations, Left =
0.444, Right = 0.441; importance for age, Left = 7.5\%, Right =
8.3\%) and the L/R pericalcarine cortex (mean RF correlations, Left =
0.267, Right = 0.315; importance for age, Left = 3.6\%, Right = 6.3\%).
In many cases, however, age was not an important variable in
high-performing models, such as that for the left accumbens area (mean
RF correlations = 0.275; relative age importances = 1.2\%).

>>>>

\subsubsection*{Gut microbial taxa are associated with specific cognitive functions} %TODO: add numbers!

Given that the development of different domains of cognitive function
(eg gross motor development vs visual perception) may develop at different times
and be affected by different molecular signals,
we sought to determine whether the associations of microbial taxa
with cognitive function are general across all cognitive subdomains,
or specific.
To accomplish this, we assessed individual subscales of the  
Mullen Scales of Early Learning (MSEL - \textit{see Methods}). 
Taxonomic profiles coupled with demographics exhibited predictive potential towards three probed subscales (Figure 4A), 
namely Expressive Language (EL; test-set RMSE XXX, test-set correlation YYYY), 
Gross Motor (GM; test-set RMSE XXX, test-set correlation YYYY) and 
Visual Reception (VR; test-set RMSE XXX, test-set correlation YYYY).
Between the subscales, not only the predictability (measured by the test-set figures of merit) is different,
but also the relative importances of several taxa (Figure 4B).
Although certain species such as \textit{B. longum}, \textit{F. prausznitzii} 
have similar relative importance across all subscales, others are substantially different.
Notably, \textit{Bifidobacterium pseudocatenulatum}, \textit{B. wexlerae} and \textit{E. eligens} are
substantially more important for the prediction of EL, while
\textit{R. faecis}, \textit{S. salivarius} and \textit{F. saccharivorans} are
distinctly predictive of GM, and
\textit{Clostridium innocuum}, \textit{B. vulgatus} 
stand out on models targeting VR.

@@

Given that individual microbes could be associated with particular domains
of cognitive function and not others, if there are causal effects of microbial metabolism on cognitive
function, they might be reflected in changes in neuroanatomy. We again
used RF modeling to associate gut taxonomic
profiles with individual brain regions identified in MRI scans
(children over 18 months old, N = 124).
Some brain regions were more readily
predicted by RF models trained on microbial taxa (Table 4, Table S2), in particular those that were highly correlated with age.
These included the L/R lingual gyrus (mean RF correlations, Left =
0.444, Right = 0.441; importance for age, Left = 7.5\%, Right =
8.3\%) and the L/R pericalcarine cortex (mean RF correlations, Left =
0.267, Right = 0.315; importance for age, Left = 3.6\%, Right = 6.3\%).
In many cases, however, age was not an important variable in
high-performing models, such as that for the left accumbens area (mean
RF correlations = 0.275; relative age importances = 1.2\%).
